Wall-to-wall mapping of peat depth from Lidar terrain and airborne radiometrics in Norwegian landscapes

### 2024-11-26 ms/ms.Rmd
add best model's mean error
ms: Intro - add ref

Although theory suggests that one meter of peat may fully attenuate radiometric signal [@beamishGammaRayAttenuation2013; @reinhardtGammaraySpectrometryVersatile2019], empirical investigations show that the association between peat depth and radioactivity can extend beyond the first meter [@gatisMappingUplandPeat2019; @kogantiMappingPeatDepth2023]. 
Radiometric data are increasingly available over large areas, presenting a big opportunity insofar as they are predictive. 
Airborne surveys are used for various applications, and some countries have high spatial coverage of such data  [@minasnyDigitalMappingPeatlands2019; @baranwalAirborneGeophysicalSurveys2020].

@@

The best models at both sites strongly underpredicted the deepest peats (Fig. \@ref(fig:calPlots)).

>>>>

Although theory suggests that one meter of peat may fully attenuate radiometric signal [@beamishGammaRayAttenuation2013; @reinhardtGammaraySpectrometryVersatile2019], empirical investigations show that the association between peat depth and radioactivity can extend beyond the first meter [@keaneySpatialStatisticsEstimate2013; @gatisMappingUplandPeat2019; @kogantiMappingPeatDepth2023]. 
Radiometric data are increasingly available over large areas, presenting a big opportunity insofar as they are predictive. 
Airborne surveys are used for various applications, and some countries have high spatial coverage of such data  [@minasnyDigitalMappingPeatlands2019; @baranwalAirborneGeophysicalSurveys2020].

@@

The best models at both sites overpredicted shallow peats and strongly underpredicted very deep peats (Fig. \@ref(fig:calPlots)).
The mean error (bias) of these models was \unit{10\,cm} at Skrimfjella and \unit{-4\,cm} at Ørskogfjellet.

### 2024-11-26 ms/ms.tex
add best model's mean error
ms: Intro - add ref

Although theory suggests that one meter of peat may fully attenuate radiometric signal \citep{beamishGammaRayAttenuation2013, reinhardtGammaraySpectrometryVersatile2019}, empirical investigations show that the association between peat depth and radioactivity can extend beyond the first meter \citep{gatisMappingUplandPeat2019, kogantiMappingPeatDepth2023}.
Radiometric data are increasingly available over large areas, presenting a big opportunity insofar as they are predictive.
Airborne surveys are used for various applications, and some countries have high spatial coverage of such data \citep{minasnyDigitalMappingPeatlands2019, baranwalAirborneGeophysicalSurveys2020}.

@@

The best models at both sites strongly underpredicted the deepest peats (Fig. \ref{fig:calPlots}).

>>>>

Although theory suggests that one meter of peat may fully attenuate radiometric signal \citep{beamishGammaRayAttenuation2013, reinhardtGammaraySpectrometryVersatile2019}, empirical investigations show that the association between peat depth and radioactivity can extend beyond the first meter \citep{keaneySpatialStatisticsEstimate2013, gatisMappingUplandPeat2019, kogantiMappingPeatDepth2023}.
Radiometric data are increasingly available over large areas, presenting a big opportunity insofar as they are predictive.
Airborne surveys are used for various applications, and some countries have high spatial coverage of such data \citep{minasnyDigitalMappingPeatlands2019, baranwalAirborneGeophysicalSurveys2020}.

@@

The best models at both sites overpredicted shallow peats and strongly underpredicted very deep peats (Fig. \ref{fig:calPlots}).
The mean error (bias) of these models was \unit{10\,cm} at Skrimfjella and \unit{-4\,cm} at Ørskogfjellet.
